### 2023-06-08 popl24/minicat.tex
checkpoint

\section{The $\minifed$ Language}

The $\minifed$ language provides a simple model of synchronous
protocols between a federation of \emph{clients} exchanging values in
the binary field. We will identify clients by natural numbers, and
federations- finite sets of clients- are always given statically.
As we will see, our threat model assumes a partition of the federation
into \emph{honest} $H$ and \emph{corrupt} $C$ subsets.

We model probabilistic programming via the \emph{random tape}
approach. That is, we will assume that programs can make reference to
values chosen from a uniform random distribution- coin ``flips''- via
values arbitrarily assigned in the initial program memory.  Programs
execute deterministically given the random tape. The random tape
formulation supports our automated analysis as we discuss in
section

\subsection{Syntax} The syntax of $\minifed$, defined in
Figure \ref{fig-minifed-syntax}, includes a standard boolean algebra
with $\eand$, $\eor$, $\exor$, and $\enot$ as primitives. In addition
programs have different kinds of variables, including \emph{secrets}
$\secret{\cid}{w}$, \emph{flips} $\flip{\cid}{w}$, and \emph{views}
$\view{\cid}{w}$.  Each of these variables are indexed by an ``owner''
client $\cid$ and distinguishing string $w$. So for example,
$\secret{\ttt{1}}{\ttt{foo}}$ is client $\ttt{1}$'s secret called
$\ttt{foo}$. All clients can make reference to the \emph{oracle}
through variables $\oracle{w}$- this is a form of shared randomness
that is standard in the MPC setting \cite{XXX}.  Each client can only
compute on their own variables in \emph{expressions} $\be$, and share
values with other clients by assignment to their views as in
$\eassign{\view{\cid}{w}}{\be}$.  A \emph{protocol} $\prog$ is a
possibly empty sequence of view assignments. We will generally omit
$\varnothing$ from example code, writing $\instr_1;\ldots;\instr_n$
instead of $\instr_1;\ldots;\instr_n;\varnothing$, and abusing notation we will
write $\prog_1;\prog_2$ to denote the concatenation of $\prog_1$
and $\prog_2$.

We let $X$ range over sets of variables and $S,V,F$ to range over sets
of secrets, views, and flips respectively. Given a program $\prog$, we
write $\iov(\prog)$ to denote $S \cup V$ where $S$ is the set of
secret variables in $\prog$ and $V$ is the set of views in $\prog$,
and we write $\flips(\prog)$ to denote the set $F$ of flip variables
in $\prog$. For any set of variables $X$ and parties $P$, we write
$X_P$ to denote the subset of $X$ owned by any party in $P$, and we
write $X_H$ and $X_C$ to denote the subsets belonging to honest and
corrupt parties, respectively.

Expressions forms include the convenience
$\select{\be_1}{\be_2}{\be_3}$ which is essentially a conditional
expression with:
$$
\select{\be_1}{\be_2}{\be_3} \equiv (\be_1\ \eand\ \be_2)\ \eor\ (\enot\ \be_1\ \eand\ \be_3)
$$
We also include \emph{oblivious transfer} $\OT{\be_1}{\be_2}{\be_3}$ as a primitive,
with semantics similar to $\ttt{select}$ but with important nuances
related to communication between clients as we discuss more below.
As we will demonstrate, it is not necessary to include $\ttt{OT}$
as a primitive since we can implement it in a provably secure
fashion. However its inclusion simplifies our presentation and is a useful
convenience.

\subsection{Semantics}

\begin{fpfig}[t]{$\minifed$ source code syntax.}{fig-minifed-syntax}
$$
\begin{array}{rcl@{\hspace{8mm}}r}
b &\in& \{ \etrue, \efalse \} \\
w &\in& \mathrm{String} \\ 
\cid &\in& \mathrm{Clients} \subset  \mathbb{N} \\[2mm]%\qquad  (\mathrm{Clients} \subset \mathbb{N} \text{\ in\ this\ presentation})\\[2mm]
\bop &\in& \{ \eand, \eor, \exor \} \\[2mm]
\be &::=& b \mid \flip{\cid}{w} \mid \secret{\cid}{w} \mid \view{\cid}{w} \mid \oracle{w} \mid & \textit{boolean expressions}\\
& &  \enot\ \be \mid \be\ \bop\ \be \mid \select{\be}{\be}{\be} \mid \OT{\be}{\be}{\be} \\[2mm]
\instr &::=& \eassign{\view{\cid}{w}}{\be} & \textit{view assignments} \\[2mm]
\prog &::=& \varnothing \mid \instr; \prog & \textit{protocols}
\end{array}
$$ 
\end{fpfig}

\emph{Memories} are fundamental to the semantics of $\fedcat$ and
provide the random tape and secret inputs to protocols, and record
view assignments. Memories $\store$ are finite (partial) mapping from
variables to binary values $\beta \in \{0,1\}$. The \emph{domain} of a
memory is written $\dom(\store)$ and is the finite set of variables on
which the memory is defined. We write $\store\{ x \mapsto \beta\}$ for
$x\not\in\dom(\store)$ to denote the memory $\store'$ such that
$\store'(x) = \beta$ and otherwise $\store'(y) = \store(y)$ for all $y \in
\dom(\store)$.

Given a set of variables $X$, we write $\store_X$ to denoted the
memory $\store$ restricted to the domain $X$, and and we define
$\mems(X)$ as the set of all memories with domain $X$:
$$
\mems(X) \defeq \{ \store \mid \dom(\store) = X \}
$$
Thus, given a protocol $\prog$, the set of all random tapes for
$\prog$ is $\mems(\flips(\prog))$. We let $\stores$ range
over sets of memories.

Given a variable-free boolean expression $\be$, we write $\cod{\be}$
to denote the standard interpretation of $\be$ in the binary field.
With the introduction of variables to expressions, we have two
concerns. First, we need to interpret variables with respect to a
specific memory, and second, we need to ensure that variables are used
``legally''. That is, since expressions define local computation, all
variables used in an expression must belong to the same client.  Thus,
we denote interpretation of expressions $\be$ possibly containing
variables as $\lcod{\store,\be}{\cid}$, where $\store$ associates
variables with values and all variables must be owned by client
$\cid$. This is defined in Figure \ref{fig-minifed-interp}.

\begin{fpfig}[t]{$\minifed$ expression interpretation.}{fig-minifed-interp}
\begin{eqnarray*}
\lcod{\store, \etrue}{\cid} &=& 1\\
\lcod{\store, \efalse}{\cid} &=& 0\\
\lcod{\store, \flip{\cid}{w}}{\cid} &=& \store(\flip{\cid}{w})\\
\lcod{\store, \secret{\cid}{w}}{\cid} &=& \store(\secret{\cid}{w})\\
\lcod{\store, \view{\cid}{w}}{\cid} &=& \store(\view{\cid}{w})\\
\lcod{\store, \oracle{w}}{\cid} &=& \store(\oracle{w})\\
\lcod{\store, \enot\ \be}{\cid} &=& \cod{\enot\ \lcod{\store,\be}{\cid}}\\
\lcod{\store, \be_1\ \mathit{binop}\ \be_2}{\cid} &=&
    \cod{\lcod{\store,\be_1}{\cid}\ \mathit{binop}\ \lcod{\store,\be_2}{\cid}}\\
\lcod{\store, \select{\be_1}{\be_2}{\be_3}}{\cid} &=&
             \begin{cases}
                \lcod{\store,\be_2}{\cid} & \text{if\ } \lcod{\store,\be_1}{\cid}\\
                \lcod{\store,\be_3}{\cid} & \text{if\ } \neg\lcod{\store,\be_1}{\cid}
             \end{cases}%\\
%\lcod{\store, \OT{\be_1}{\be_2}{\be_3}}{\cid} &=&
%             \begin{cases}
%                \lcod{\store,\be_2}{\cid'} & \text{if\ } \lcod{\store,\be_1}{\cid}\\
%                \lcod{\store,\be_3}{\cid'} & \text{if\ } \neg\lcod{\store,\be_1}{\cid}
%             \end{cases}
\end{eqnarray*}
\end{fpfig}

Evaluation of configurations is then defined via a small-step reduction relation $\redx$.
This is defined in the obvious manner for view assignments other than through
$\ttt{OT}$- however note that reduction requires that views are never reassigned. 
\begin{mathpar}
  (\store, \eassign{\view{\cid}{w}}{\be};\prog) \redx (\extend{\store}{\view{\cid}{w}}{\lcod{\store,\be}{\cid'}}, \prog)
\end{mathpar}
For view assignments through $\ttt{OT}$, we define a different reduction rule that
captures the appropriate semantics- that is, the selection bit is \emph{not} communicated
to the sender (through a view assignment), and only the selected bit is sent to the receiver:
\begin{mathpar}
  \inferrule
  {\beta = \text{if\ } \lcod{\store,\be_1}{\cid}  \text{\ then\ } \lcod{\store,\be_2}{\cid'} \text{\ else\ } \lcod{\store,\be_3}{\cid'}}
      {(\store, \eassign{\view{\cid}{w}}{\OT{\be_1}{\be_2}{\be_3}};\prog) \redx (\extend{\store}{\view{\cid}{w}}{\beta}, \prog)}
\end{mathpar}
We define $\redxs$ as the reflexive, transitive closure of $\redx$.

Given $\prog$ with $\iov(\prog) = S \cup V$ and $\flips(\prog) = F$,
any execution of $\prog$ is assumed to be an evaluation of a
configuration $\config{\store}{\prog}$ where $\store \in \mems(S \cup
F)$- that is, with secrets and the random tape as inputs- and where
$\config{\store}{\prog} \redxs \config{\store'}{\varnothing}$ it is
the case that $\store' \in \mems(S \cup F \cup V)$.

\subsection{Examples}

Here we introduce some examples to illustrate $\minifed$ and how it
can be used to model secure communication protocols for values in the
binary field. Most of these are translations (into $\minifed$) of
object examples in previous related work \cite{XXX}. We will return to
these as running examples throughout the paper to illustrate various
concepts and advantages of our methods.
\begin{example}[One-Time Pad]
  \label{example-otp}
As observed in \cite{XXX} and elsewhere, coin flips with $\exor$ can
be used to model one-time pad encryption in the binary field. On the
basis of this we can model secure communication between over an
insecure channel using symmetric key encryption. In the following
program, we assume $H = \{ 1,2 \}$ and $C = \{ 0 \}$. Client 1 first
shares the randomly generated ``key'' $\fx{1}{0}$ with client 2
through view $\view{2}{0}$. Then, client 1 encrypts its secret
$\sx{1}{0}$ and sends it to party 0, which relays the ciphertext in
$\vx{0}{0}$ to client 2. Finally, client 2 decrypts the the ciphertext
and stores the plaintext in $\vx{2}{2}$.
\begin{verbatimtab}
v[2,0] := f[1,0];
v[0,0] := f[1,0] xor s[1,0];
v[2,1] := v[0,0];
v[2,2] := v[2,1] xor v[2,0]
\end{verbatimtab}
Intuitively, this protocol is \emph{correct} because $\vx{2}{2}$ is
exactly correlated with the secret $\sx{1}{0}$, and it is
\emph{secure} because the corrupt view $\vx{0}{0}$ in isolation is
probabilistically independent of the secret $\sx{1}{0}$.
\end{example}

\begin{example}[Lambda-obliv]
  \label{example-lambda-obliv}

  \begin{verbatimtab}
v[0,0] := select(s[1,0], flip[1,0], flip[1,1] );
v[0,1] := flip[1,0]
\end{verbatimtab}

\begin{verbatimtab}
v[0,0] := select(s[1,0],select(flip[1,0], flip[1,0], flip[1,1]),flip[1,2])
\end{verbatimtab}
\end{example}

\begin{example}[Oblivious Transfer]
  \label{example-lambda-OT}

\begin{verbatimtab}
v[2, rd] := select[flip[2, d], H[r1], H[r0]];
v[1, e] := s[2, c] xor flip[2, d];
v[2, f0] := s[1, m0] xor select[v[1, e], H[r1], H[r0]];
v[2, f1] := s[1, m1] xor select[not v[1, e], H[r1], H[r0]];
v[2, mc] := (select[s[2, c], v[2, f1], v[2, f0]] xor v[2, rd])
\end{verbatimtab}
\end{example}

\begin{example}[Additive Secret Sharing]
    \label{example-lambda-he}
Another 
example of an MPC protocol is additive secret sharing, where $k$
parties split up their secrets into \emph{shares}, with the property
that all $k$ shares are required to reconstruct the secret and any
fewer reveals nothing. In $\fedcat$, we can use $\ttt{xor}$, again, to
generate shares. Here, party 1 generates 3 shares of the secret
$\sx{1}{1}$- one is given to party 2 through view $\vx{2}{s1}$, one is
given to party 3 through view $\vx{3}{s1}$, and party 1 keeps
$\fx{1}{s1}$ for itself:
\begin{verbatimtab}
v[2,s1] := flip[1,1] xor flip[1,s1] xor s[1,1];
v[3,s1] := flip[1,1];
\end{verbatimtab}
Note that $\vx{2}{s1}$ and $\vx{3}{s1}$ are both in uniform
random distributions.

Another property of $\ttt{xor}$ is that it is additively homomorphic,
while functioning as addition in the boolean algebra (i.e., the binary
field). That is, given:
\begin{mathpar}
  x_1\ \ttt{xor}\ y_1 = z_1

  x_2\ \ttt{xor}\ y_2 = z_2
\end{mathpar}
then we have:
$$
(x_1\ \ttt{xor}\ x_2)\ \ttt{xor}\ (y_1\ \ttt{xor}\ y_2) = z_1\ \ttt{xor}\ z_2
$$
Thus, we can obtain secret 3-party summation (xor) by continuing
the protocol begun above- parties 2 and 3 can share their secrets
in the same fashion:
\begin{verbatimtab}
v[1,s2] := flip[2,1] xor flip[2,s2] xor s[2,1];
v[3,s2] := flip[2,1];

v[1,s3] := flip[3,1] xor flip[3,s3] xor s[3,1];
v[2,s3] := flip[3,1];
\end{verbatimtab}
Then, assuming that party 0 is the ``public party'', each party sums
its own shares, and then the sum of sum of shares is revealed
in the output view $\vx{0}{output}$. By additive homomorphism of
$\ttt{xor}$, this output is the sum of the 3 secrets, but
the protocol reveals nothing to the adversary, even if one
(or even two, in fact) of $\{1,2,3\}$ are corrupt. 
\begin{verbatimtab}
v[0,ss1] := v[1,s2] xor v[1,s3] xor flip[1,s1];
v[0,ss2] := v[2,s1] xor v[2,s3] xor flip[2,s2];
v[0,ss3] := v[3,s1] xor v[3,s2] xor flip[3,s3];

v[0,output] := v[0,ss1] xor v[0,ss2] xor v[0,ss3]
\end{verbatimtab}

Because this protocol correctly implements 3-party addition in the
binary field, it can reveal information to the adversary. For example,
suppose that $1 \in C$, and $\sx{1}{1}$ is $1$, and after running the
protocol suppose that $\vx{0}{output}$ is 0. Then the adversary knows
that 2 and 3's secrets are either 1 and 0, or 0 and 1, with 50/50
probability, and definitely neither 0 and 0, nor 1 and 1. However, the
protocol does not reveal anything more through adversarial views than
what the output reveals.
\end{example}