### 2023-07-04 popl24/main.tex
checkpoint

\documentclass[acmsmall,screen,review]{acmart}

\usepackage{amsmath}
\usepackage{amstext}
\usepackage{fp-frame}
\usepackage[latin1]{inputenc}
\usepackage{mathpartir}
\usepackage{fancyvrb}
\usepackage{moreverb}
\usepackage{stmaryrd}
\usepackage{enumerate}

\input{macros}

\begin{document}

\title{Automating Verification of MPC Security: $\metaprot$ to $\minifed$}

\author{Author names withheld for double-blind reviewing.}

\begin{abstract}
  We formulate a sound hyperproperty of MPC passive security and show how this
  supports automated and semi-automated verification.
\end{abstract}

%\maketitle

%\input{intro}

%\input{minicat}

%\input{pmf}

%\input{hyperprop}

%\input{bruteforce}

%\input{metalang}

%\input{composition}

\input{scratch}

\bibliographystyle{ACM-Reference-Format}
\bibliography{logic-bibliography,secure-computation-bibliography}

\end{document}
\endinput