### 2023-07-10 popl24/gmw.tex
checkpoint

\begin{fpfig}[t]{2-Party GMW Circuit Library with And gate.}{fig-gmw}
{\footnotesize
  \begin{verbatimtab}
    encodegmw(in1,in2) {
      v[1, in1 || "out"] := s[1,in1] xor flip[1,in1];
      v[2, in1 || "out"] := flip[1,in1];
      v[1, in2 || "out"] := s[2,in2] xor flip[2,in2];
      v[2, in2 || "out"] := flip[2,in2]
      { shares1 = { c1 = v[1, in1 || "out"]; c2 = v[2, in1 || "out"] };
        shares2 = { c1 = v[1, in2 || "out"]; c2 = v[2, in2 || "out"]} } 
    }

    andtablegmw(b1, b2) {
      let r11 = (b1 xor true) and (b2 xor true) in
      let r10 = (b1 xor true) and (b2 xor false) in
      let r01 = (b1 xor false) and (b2 xor true) in
      let r00 = (bl xor false) and (b2 xor false) in
      { v1 = r11; v2 = r10; v3 = r01; v4 = r00 }
    }

    andgmw(g, shares1, shares2) {
      let r = flip[1,g || ".r"] in
      let table = andtablegmw(shares1.c1, shares2.c1) in
      let r11 =  r xor table.v1 in
      let r10 =  r xor table.v2 in
      let r01 =  r xor table.v3 in
      let r00 =  r xor table.v4 in
      v[2,g || "out"] := OT4(shares1.c2, shares2.c2, r11, r10, r01, r00);
      v[1,g || "out"] := r;
      { c1 = v[1,g || "out"]; c2 = v[2,g || "out"]}
    }

    decodegmw(shares) { v[0,"output"] := shares.c1 xor shares.c2 }
  \end{verbatimtab}
}
\end{fpfig}