### 2023-07-11 popl24/main.tex

\documentclass[acmsmall,screen,review]{acmart}

\usepackage{amsmath}
\usepackage{amstext}
\usepackage{fp-frame}
\usepackage[latin1]{inputenc}
\usepackage{mathpartir}
\usepackage{fancyvrb}
\usepackage{moreverb}
\usepackage{stmaryrd}
\usepackage{enumerate}
\usepackage{thmtools,thm-restate}
\usepackage{comment}

\input{macros}

\begin{document}

\title{Automating Verification of MPC Security: $\metaprot$ to $\minifed$}

\author{Author names withheld for double-blind reviewing}

\begin{abstract}
Secure Multi-Party Computation (MPC) protocols support data privacy in
important modern distributed applications. Security for MPC protocols
is superficially similar to probabilistic noninterference, but differs
in a subtle but fundamental way, and approaches for verifying
noninterference cannot naturally extend to MPC security.  Currently,
proof methods for MPC protocols are well-studied but manual and thus
tedious and error-prone, and are also non-standardized and unfamiliar
to most PL theorists.  Our goal is to leverage connections between the
security model of MPC and trace-based hyperproperties to obtain
automated proof methods for MPC protocol development.  We develop a
language model with a tightly coupled notion of probabilistic program
distributions, as a foundation for fully and partially automated
verification of passive MPC security in protocols including
arbitrarily large YGC and GMW circuits.
\end{abstract}

\maketitle

\input{intro}

\input{minicat}

\input{metalang}

\input{pmf}

\input{hyperprop}

\input{automation}

\input{composition}

%% This is scratch pad to develop ideas before integration.
%\input{scratch}

\bibliographystyle{ACM-Reference-Format}
\bibliography{logic-bibliography,secure-computation-bibliography}

\appendix

\input{examples}

%\input{types}

\input{ygc}

\input{bruteforce}

\input{proofs}

\end{document}
\endinput