### 2024-04-17 revision-1/model.tex

\section{Security Model}

\subsection{Passive Security}

The simulator is represented by a probabilistic algorithm $\SIM_C$,
aka a \emph{simulation}, that is parameterized by corrupt inputs and
the output of an ideal functionality, and that returns a set of
adversarial views (as a memory) with some probability. Given
corrupt inputs $\store$ and ideal functionality output $v$,  
we write
$
\prob(\SIM(\store,v) = \store')
$
to denote the probability that $\SIM(\store,v)$
returns corrupt views $\store'$ as a result. We can then define the
probability distribution of corrupt views reconstructed
by the simulator as follows:
\begin{definition}
  Given $C$, $\store$, and $v$, we write $\dist(\SIM(\store,v))$ to
  denote the distribution of corrupt views reconstructed by the
  simulation, where for
  all $\store' \in \mems(V)$:
  $$
  \dist(\SIM(\store,v))(\store')\ \defeq\ \prob(\SIM(\store,v) = \store') 
  $$
\end{definition}

Then we can define passive security in the real/ideal
model as follows. 
\begin{definition}[Passive Security]
  Assume given a program $\prog$ that correctly implements an ideal
  functionality $\idealf$, with $\iov(\prog) = (S,M,O)$.  Then $\prog$
  is \emph{passive secure in the simulator model} iff for all
  partitions of the federation into honest and corrupt sets $H$ and $C$
  with $|C| < |H|$ and for all $\store \in \mems(S)$ there exists a
  simulation $\SIM$ such that:
  $$
  \dist(\SIM(\store_{S_C},\idealf(\store))) = \condd{\progtt(\prog)}{M_C}{\store}
  $$
\end{definition}

\subsection{Malicious Security}

$$
\begin{array}{rclr}
  (\store, \eassign{\mesg{w}}{\cid_1}{\be}{\cid_2};\prog) &\aredx&
  (\extend{\store}{\mesg{w}_{\cid_1}}{\lcod{\store,\be}{\cid_2}}, \prog) & \cid_2 \in H\\
  (\store, \eassign{\mesg{w}}{\cid_1}{\be}{\cid_2};\prog) &\aredx&
  (\extend{\store}{\mesg{w}_{\cid_1}}{\lcod{\arewrite(\store_C,\be)}{\cid_2}}, \prog) & \cid_2 \in C
\end{array}
$$

\begin{definition}[Corrupt Inputs, Honest Outputs]
  Given a program $\prog$ with $\iov(\prog) = (S,M,O)$ , define $\cinputs$ as the
  messages in $M$ sent from corrupt to honest parties:
  $$
  \cinputs = \{\ \elab{\mesg{w}}{\cid}\ \mid\  \elab{\mesg{w}}{\cid} \in M \wedge \eassign{\mesg{w}}{\cid}{\be}{\cid'} \in \prog
  \wedge \cid \in H \wedge \cid' \in C \ \} 
  $$
  and similarly define $\houtputs$ as the messages in $M$ sent from honest to corrupt parties.
  %Define also $(\afilter\ \prog)$ as $\prog$ with all instructions of the form $\eassign{\mesg{w}}{\cid}{\be}{\cid'}$ removed
  %where $\cid \in H \wedge \cid' \in C$.
\end{definition}

$$
\dist(\SIM(\store_{S_H})) = \condd{\progtt(\prog)}{\houtputs \cup O_H}{\store_{S_H}}
$$