### 2024-05-17 ppdp24/intro.tex
checkpoint

\section{Introduction}

Secure Multi-Party Computation (MPC) protocols support data privacy in
important modern, distributed applications such as privacy-preserving
machine learning \cite{li2021privacy, knott2021crypten,
  koch2020privacy, liu2020privacy} and Zero-Knowledge proofs in
blockchains \cite{ishai2009zero, lu2019honeybadgermpc,
  gao2022symmeproof, tomaz2020preserving}. The security semantics of
MPC include both confidentiality and integrity properties incorporated
into models such as real/ideal (aka simulator) security and universal
composability (UC), developed primarily by the cryptography community
\cite{XXX}.  Related proof methods are well-studied \cite{Lindell2017}
but mostly manual. Somewhat independently, a significant
body of work in programming languages has focused on definition and
enforcement of confidentiality and integrity \emph{hyperproperties}
\cite{10.5555/1891823.1891830} such as noninterference and delimited
release \cite{XXX}. Following a tradition of connecting
cryptographic and PL-based security models \cite{XXX}, recent work has
also recognized connections between MPC security models and
hyperpoperties of, e.g., noninterference \cite{8429300}, and even
leveraged these connections to enforce MPC security through mechanisms
such as security types \cite{XXX}. Major benefits of this connection
in an MPC setting include better language abstractions for defining
protocols and for mechanisation and even automation of security proofs.
The goal of this paper is to develop a PL model for defining a variety of
low-level probabilistic MPC protocols, to formulate a collection
of confidentiality and integrity hyperproperties for our model
model with familiar information flow analogs, and to show how these
properties can be leveraged for improved proof automation.

The distinction between high- and low-level languages for MPC is
important. High-level languages such as Wysteria \cite{XXX} and
Viaduct \cite{XXX} are designed to provide effective programming of
full applications. These language designs incorporate sophisticated
verified compilation techniques such as orchestration \cite{XXX} to
guarantee high-level security properties, and they rely on
\emph{libraries} of low-level MPC protocols, such as binary and
arithmetic circuits. These low-level protocols encapsulate
abstractions such as secret sharing and semi-homomorphic encrypytion,
and rely on probabilistic progamming methods. So, low-level MPC
programming and protocol verification is a distinct challenge and both
critical to the general challenge of PL design for MPC and
complementary to high-level language design.

The connection between information flow hyperproperties and
MPC security is also complicated especially at a low level.
MPC protocols involve communication between a group of distributed
clients called a \emph{federation} that collaboratively compute
and publish the result of some known \emph{ideal functionality}
$\idealf$, maintaining confidentiality of inputs to $\idealf$ without
the use of a trusted third party. However, since the outputs of
$\idealf$ are public, some information about inputs is inevitably
leaked. Thus, the ideal
functionality establishes a declassification policy \cite{XXX}, which
is more difficult to enforce than pure noninterference.  And
subtleties of, e.g., semi-homomorphic encryption are central to both
confidentiality and integrity properties of protocols and similarly
difficult to track with coarse-grained security types alone.

Nevertheless, as previous authors have observed \cite{XXX}, low-level
protocol design patterns such as secret sharing and circuit gate
structure have compositional properties that can be abstracted and
leveraged in larger proof contexts. We contribute to this line of work
by developing an automated verification technique for subprotocols and
show how it can be integrated as a tactic in a larger security proof.

\compfig

\subsection{Related Work}
\label{section-related-work}

Our main focus is on PL design and automated and semi-automated
reasoning about security properties of low-level MPC protocols.  We
consider related systems in several dimensions, including
whether they are aimed at low-level design with probabilistic features,
whether they support reasoning about conditional probabilities which
are central to real/ideal security, whether they consider MPC
security through the lens of hyperproperties, and whether they consider
passive and/or malicious security models. We summarize this comparison
in Figure \ref{fig-comp}, with the caveat that works vary in the degree
of development in each dimension.

As mentioned above, several high-level languages have been developed
for writing MPC applications, 
and frequently exploit the connection between hyperproperties and MPC
security. Previous work on analysis for the SecreC language
\cite{almeida2018enforcing,10.1145/2637113.2637119} is concerned with
properties of complex MPC circuits, in particular a user-friendly
specification and automated enforcement of declassification bounds in
programs that use MPC in subprograms. The Wys$^\star$ language
\cite{wysstar}, based on Wysteria \cite{rastogi2014wysteria}, has
similar goals and includes a trace-based semantics for reasoning about
the interactions of MPC protocols. Their compiler also guarantees that
underlying multi-threaded protocols enforce the single-threaded source
language semantics. These two lines of work were focused on passive
security. The Viaduct language \cite{XXX} has a well-developed
information flow type system that automatically enforces both
confidentiality and integrity through hyperproperties such as robust
declassification, in addition to rigorous compilation guarantees
through orchestration \cite{XXX}. However, these high level
languages lack probabilistic features and other abstractions of
low-level protocols, the implementation and security of which
are typically assumed as a selection of library compnonents.

Various related low-level languages with probabilistic features have
also been developed. The $\lambda_{\mathrm{obliv}}$ language
\cite{darais2019language} uses a type system is used to automatically
enforce so-called probabilistic trace obliviousness.  But similar to
previous work on oblivious data structures \cite{10.1145/3498713},
obliviousness is related to pure noninterference, not the relaxed form
related to passive MPC security. The Haskell-based security type system in
\cite{XXX} enforces a version of conditioned noninterence that is
sound for passive security in a probabilisitic low-level setting, but
they do not consider malicious security. And properties of real/ideal
passive and malicious security for a probabilistic language have been
formulated in EasyCrypt \cite{8429300}-- though their proof methods,
while mechanized, are fully manual, and their formulation of malicious
security is not as clearly related to robust declassification as the
one we present in Section \ref{section-hyperprop}.

Program logics for probabilistic languages and specifically reasoning
about properties such as joint probabilistic independence is also
important related work. Probabilistic Separation Logic (PSL)
\cite{barthe2019probabilistic} develops a logical framework for
reasoning about probabilistic independence (aka separation) in
programs, and they consider several (hyper)properties, such as perfect
secrecy of one-time-pads and indistinguishability in secret sharing,
that are critical to MPC. However, their methods are manual, and they
do not capture notions of conditional independence (separation). This
latter issue has been addressed in Lilac \cite{XXX}, but the
application of Lilac-style reasoning to MPC protocols has not
previously been explored, as we do in Section
\ref{section-example-gmw}.

Our work also shares many ideas with probabilistic programming
languages designed to perform (exact or approximate) statistical
inference~\cite{holtzen2020scaling, carpenter2017stan, wood2014new,
  bingham2019pyro, albarghouthi2017fairsquare, de2007problog,
  pfeffer2009figaro, saad2021sppl}. Our setting, however, requires
verifying properties beyond inference, including conditional
statistical independence. Recent work by Li et al.~\cite{li2023lilac} proposes a
manual approach for proving such properties, but does not provide
automation.

\subsection{Overview and Contributions}

\paragraph{Language design.} In Section \ref{section-minicat} we
develop a new probabilistic programming language $\minifed$ for
defining synchronous distributed protocols over the binary field. The
syntax and semantics provides a succinct account of \emph{views} and
(synchronous) messaging between protocol \emph{clients}. In Section
\ref{section-metalang} we define a new metaprogramming language
$\metaprot$ that dynamically generates $\minifed$ protocols. It
includes control and data structures and is able to express logical
protocol components, and enjoys a type safety result that guarantees
that generated protocols are semantically well-defined.

\paragraph{Hyperproperty formulation.} In Section \ref{section-pmf} we
formulate \emph{program distributions}, a formalism for expressing
dependencies between secrets and views, useful for expressing
characteristics of protocols and precisely quantifying allowable
information leakage in terms of protocol outputs. On this basis we can
formulate security hyperproperties, including \emph{noninterference
modulo output ($\NIMO$)} that implies passive simulator security. But
program distributions are flexible and can be used to certify other
important properties of protocol components, such as probabilistic
independence, correlation, and distributions under conditioning.

\paragraph{Fully and partially automated verification.} In Section
\ref{section-automation} we develop an automated technique, aka
certification, for automatically verifying security properties such as
$\NIMO$ and other supporting properties of protocol components.  While
certification has high complexity, a compositional approach allows
certification of program components in isolation.  This enables a
semi-automated method for large circuits where we (1) define
certification for components, (2) certify components, (3) prove
$\NIMO$ based on the certifiedness of components. We develop novel
methods partially based on separation logic
\cite{barthe2019probabilistic} to enable (3).  In Section
\ref{section-composition} we illustrate our method to obtain
semi-automated verification for the GMW and Yao's Garbled
Circuit protocols, where certification can be re-used to validate new
gate implementations.