### 2024-05-22 ppdp24/conclusion.tex
checkpoint

\section{Conclusion and Future Work}

The $\metaprot/\minifed$ language model developed in this paper, along
with the suite of hyperproperties explored, are a foundation
for defining low-level MPC protocols and exploring methods for security
verification.  There are several promising avenues for future
work. Concurrency will be important to capture common MPC idioms such
as commitment and circuit optimizations, and to consider the UC
security model \cite{evans2018pragmatic,viaduct-UC}. To support better
verification, integration of our tactics with a proof assistant is a
compelling direction that resonate with previous work \cite{8429300}.

Better automated verification is a main goal for future work.
Security types have been effectively leveraged for high-level
programs, for both passive and malicious security. This suggests that
they can be similarly applied at low levels in a way that is
complementary to high-level systems. Generalizing our automated
analysis in $\mathbb{F}_2$ to larger arithmetic fields is also a
compelling technical challenge with potential practical benefits,
e.g., application to arithmetic circuits.