### 2024-07-17 revision-2/gmw.tex

\begin{fpfig}[t]{2-Party GMW circuit library.}{fig-gmw}
{\footnotesize
  \begin{verbatimtab}
    encodegmw(in, i1, i2) {
      m[in]@i2 := (s[in] xor r[in])@i2;
      m[in]@i1 := r[in]@i2;
      m[in]
    }

    andtablegmw(b1, b2, r) {
      let r11 = r xor (b1 xor true) and (b2 xor true) in
      let r10 = r xor (b1 xor true) and (b2 xor false) in
      let r01 = r xor (b1 xor false) and (b2 xor true) in
      let r00 = r xor (bl xor false) and (b2 xor false) in
      { row1 = r11; row2 = r10; row3 = r01; row4 = r00 }
    }

    andgmw(z, x, y) {
      let r = r[z] in
      let table = andtablegmw(x,y,r) in
      m[z]@2 := OT4(x,y,table,2,1);
      m[z]@1 := r@1;
      m[z]
    }

    xorgmw(z, x, y)
    { m[z]@1 := (x xor y)@1; m[z]@2 := (x xor y)@2; m[z] }

    decodegmw(z) {
      p["1"] := z@1; p["2"] := z@2;
      out@1 := (p["1"] xor p["2"])@1;
      out@2 := (p["1"] xor p["2"])@2
    }
  \end{verbatimtab}
}
\end{fpfig}