### 2024-08-18 csf25/main.tex
checkpoint

\documentclass[conference]{IEEEtran}
\IEEEoverridecommandlockouts
% The preceding line is only needed to identify funding in the first footnote. If that is unneeded, please comment it out.

\usepackage{amsmath}
\usepackage{amstext}
\usepackage{amssymb}
\usepackage{fp-frame}
\usepackage[latin1]{inputenc}
\usepackage{mathpartir}
\usepackage{fancyvrb}
\usepackage{moreverb}
\usepackage{stmaryrd}
\usepackage{enumerate}
\usepackage{thmtools,thm-restate}
\usepackage{comment}
\usepackage{booktabs,array}
\usepackage{rotating}
\usepackage{xcolor}

\input{macros}
%\input{figures}

\begin{document}

\input{lang}

\input{smt}

\input{ctj}

\input{itj}

\input{metalang}

\input{examples}

\bibliographystyle{IEEEtran}
\bibliography{logic-bibliography,secure-computation-bibliography}

\end{document}
\endinput