### 2024-08-18 csf25/lang.tex
checkpoint

\section{$\minicat$ Syntax and Operational Semantics}

\minicatsyntaxfig

\minicatredxfig

\subsection{$\minicat$ Adversarial Semantics}

$$
    \begin{array}{rcl@{\hspace{2mm}}r}
      \prog &::=& \cdots \mid \assert{\be = \be}
    \end{array}
$$

\minicataredxfig