### 2024-10-10 esop25/frontmatter.tex
checkpoint

\titlerunning{Security Types for Low-Level MPC}

\title{SMT-Boosted Security Types for Low-Level MPC}

%\author{
%Christian Skalka
%\and
%Joe Near}

%\institute{
%University of Vermont, Burlington VT 05405, USA 
%\email{ceskalka@uvm.edu,jnear@uvm.edu}
%}

\author{Author names withheld for double-blind reviewing.}
\institute{}

\newcommand{\theabstract}{
  \begin{abstract}
  Secure Multi-Party Computation (MPC) is an important
  enabling technology for data privacy in modern distributed
  applications. We devevelop a new type theory to automatically
  enforce correctness, confidentiality, and integrity properties
  of protocols written in the \emph{Prelude/Overture} language
  framework. Judgements in the type theory are predicated on
  SMT verifications in a theory of finite fields. 
\end{abstract} }