### 2024-10-11 esop25/frontmatter.tex
checkpoint

\titlerunning{Security Types for Low-Level MPC}

\title{SMT-Boosted Security Types for Low-Level MPC}

%\author{
%Christian Skalka
%\and
%Joe Near}

%\institute{
%University of Vermont, Burlington VT 05405, USA 
%\email{ceskalka@uvm.edu,jnear@uvm.edu}
%}

\author{Author names withheld for double-blind reviewing.}
\institute{}

\newcommand{\theabstract}{
  \begin{abstract}
  Secure Multi-Party Computation (MPC) is an important enabling
  technology for data privacy in modern distributed applications. We
  devevelop a new type theory to automatically enforce correctness,
  confidentiality, and integrity properties of protocols written in
  the \emph{Prelude/Overture} language framework. Judgements in the
  type theory are predicated on SMT verifications in a theory of
  finite fields, which supports precise and efficient analysis. Our
  approach is compositional, scalable, and generalizes to arbitrary
  prime fields for data and key sizes.
\end{abstract} }