### 2025-01-29 esop25/camera-ready/conclusion.tex
checkpoint

\section{Conclusion and Future Work}

In this paper we have defined confidentiality and integrity type
systems for the $\metaprot/\minifed$ language model. These type
systems resemble previous security type systems in probabilistic
settings, but their accuracy is boosted by the integral use of SMT
verification.  Our work significantly improves automation,
scalability, and flexibility of static analysis in the
$\metaprot/\minifed$ as compared to prior work.

We see several promising avenues for future work. Concurrency will be
important to capture common MPC idioms such as commitment and circuit
optimizations and to consider the UC security model
\cite{evans2018pragmatic,viaduct-UC}. We also hope to improve our type
system to enforce the declassification bounds determined by ideal
functionalities, beyond gradual release.