### 2025-04-26 esop25/slides.tex
checkpoint

% Inbuilt themes in beamer
\documentclass{beamer}

% Theme choice:
\usetheme{CambridgeUS}
\usecolortheme{default}

\usepackage[style=verbose,backend=biber]{biblatex}
\addbibresource{secure-computation-bibliography.bib}

\usepackage{mathpartir}
\usepackage{comment}
\usepackage{fancyvrb}
\usepackage{moreverb}

% \input{slide-macros}
\input{macros}

\newcommand{\sumthree}{\prog_{\mathit{sum3}}}

% Title page details: 
\title{SMT-Boosted Security Types for Low-Level MPC}

\author{Christian Skalka and Joe Near}

\date{ESOP, May 6, 2025}

\logo{
\includegraphics[width=3cm]{UVM_Logo_Primary_Horiz_G.png}
}

\beamertemplatenavigationsymbolsempty

\begin{document}

% Title page frame
\begin{frame}
    \titlepage 
\end{frame}

% Remove logo from the next slides
\logo{}

% % Outline frame
% \begin{frame}{Outline}
%     \tableofcontents
% \end{frame}

\section{Overview}

\begin{frame}{Background}

  \cemph{Secure Multiparty Computation (MPC)} is a paradigm for
  \cemph{confidential computing}.
  \begin{itemize}
  \item Supports data privacy.
  \item Uses software-based protocols (vs., e.g., enclaves).
  \item No trusted 3rd party.
  \end{itemize}
  Many existing deployments: voting, auctions, satellite trajectories, COVID exposure notification, ...\footnote{https://mpc.cs.berkeley.edu}

  \medskip

  Various language frameworks, including MPyC, Fairplay, Viaduct.
\end{frame}

\begin{frame}{Contributions}

  Existing HLLs rely on low-level MPC protocols with correctness
  assumed.
  \begin{itemize}
  \item Distinct features, e.g., probabilistic sampling.
  \item Existing proof techniques mostly manual.
  \end{itemize}
  Our contributions:
  \begin{itemize}
  \item Language design for low-level protocol definitions.
  \item Automated type analysis for security properties, with correctness based on
    \cemph{hyperproperties}\footnote{Skalka and Near, PPDP 2024}.
  \item Our type analysis leverages an \cemph{SMT finite fields theory}\footnote{Ozdemir et. al, CAV 2023} for:
    \begin{itemize}
    \item \cemph{Correctness}
    \item \cemph{Confidentiality}
    \item \cemph{Integrity}
    \end{itemize}
  \end{itemize}

\end{frame}

\section{What's an MPC Protocol}

\begin{frame}{What's an MPC Protocol?}
  A \emph{protocol} $\prog$ implements an \emph{ideal functionality} $\idealf$,
  defined in terms of operations $(+,-,*)$ in a prime field $\mathbb{F}_p$.
  \begin{itemize}
  \item The inputs are \cemph{secret values} of \cemph{clients} in a \emph{federation}
    $\setit{1,...,j}$.
  \item Clients are distributed and communicate via \cemph{message passing}.
  \item Any client can be corrupted by the adversary, who can then see their messages.
  \item The \cemph{output} is made \emph{publicly available} by $\prog$.
  \end{itemize}
  \cemph{The output may leak secret information to an adversary} trying to
  guess secret values, since it depends on inputs.

  \medskip

  \cemph{Confidentiality} means that protocol messages don't enable the adversary
  to make better guesses than the output alone.

\end{frame}

\begin{frame}{Declassification of the Ideal Functionality}

  \begin{exampleblock}{Example: 3-Party $\fplus$ in $\mathbb{F}_2$ (where $\fplus$ is $\exor$ and $\ftimes$ is $\wedge$)}
  $$
    \idealf =
    \mathit{secret}@{\mathit{client_1}} + \mathit{secret}@{\mathit{client_2}} + 
      \mathit{secret}@{\mathit{client_3}}
  $$
  Suppose in a run of $\prog$ and $\mathit{client_3}$ is
  corrupted, with $\mathit{secret}@{\mathit{client_3}} = 0$. 
  In case the output is 1, the adversary knows that either:
  $$
  \mathit{secret}@{\mathit{client_1}} = 0 \text{\ and\ } \mathit{secret}@{\mathit{client_2}} = 1
  $$
  or:
  $$
  \mathit{secret}@{\mathit{client_1}} = 1 \text{\ and\ } \mathit{secret}@{\mathit{client_2}} = 0
  $$
  with equal joint probability of .5.

  \medskip

  \cemph{Messages sent to $\mathit{client_3}$ during
    a run of $\prog$ should not
    allow the adversary better guesses.}
  \end{exampleblock}

\end{frame}

\begin{frame}{Protocol Implementation in Overture}

  Low-level, probabilistic protocols are defined in terms of basic variable types
  (identifiers $x$ distinguish them):
  \begin{itemize}
  \item $\secret{x}$: secret values in $\mathbb{F}_p$.
  \item $\mesg{x}$: messages received during the protocol.
  \item $\flip{x}$: uniformly random samples from $\mathbb{F}_p$ (implemented with
    a random tape semantics).
  \end{itemize}
  We localize computation to \emph{clients}, and any information
  exchanged between clients must be through messaging.
  %\begin{exampleblock}{Toy Example:}
  %  $$
  %  \begin{array}{rcl@{\qquad}l}
  %    \mx{x}{1} &:=& \elab{(\secret{x} - \flip{x})}{2}  & \textit{(unicast)}\\
  %    \mx{y}{*} &:=& \elab{(\flip{y} + \mesg{x})}{1}  & \textit{(broadcast)}
  %  \end{array}
  %  $$ 
  %\end{exampleblock}
\end{frame}

\section{Example: Additive Sharing}
\begin{frame}{Example: Additive Secret Shares}

  To implement 3-party summation in any $\mathbb{F}_p$, we can use \emph{additive sharing}.
  Each party first generates \emph{reconstructive shares}.

  \begin{block}{Reconstructive Sharing: Protocol $\prog_{\mathit{sum3}}$}
    $$
    \begin{array}{lll}
      \elab{\mesg{s1}}{2} &:=& \elab{(\secret{1} \fminus \locflip \fminus \flip{x})}{1} \\ 
      \elab{\mesg{s1}}{3} &:=& \elab{\flip{x}}{1} \\ 
      \elab{\mesg{s2}}{1} &:=& \elab{(\secret{2} \fminus \locflip \fminus \flip{x})}{2} \\ 
      \elab{\mesg{s2}}{3} &:=& \elab{\flip{x}}{2} \\ 
      \elab{\mesg{s3}}{1} &:=& \elab{(\secret{3} \fminus \locflip \fminus \flip{x})}{3} \\ 
      \elab{\mesg{s3}}{2} &:=& \elab{\flip{x}}{3}
    \end{array}
    $$
  \end{block}
  $\elab{\mesg{s1}}{2}, \elab{\mesg{s1}}{3}, \elab{\locflip}{1}$ are the \cemph{secret shares} of $\sx{1}{1}$,
    etc.
  \begin{itemize}
  \item \emph{algebraically} $\elab{\locflip}{1} + \elab{\mesg{s1}}{2} + \elab{\mesg{s1}}{3} \eop \sx{1}{1}$.
  \item \emph{probabilistically} any 2 out of three shares are in a uniform random distribution.
  \end{itemize}

\end{frame}

\begin{frame}{Additive Homomorphism of Shares}

  \begin{block}{Additive Homomorphism: Sum-of-Shares}     
  To sum the input secrets \emph{in secret} extend $\prog_{\mathit{sum3}}$:
    $$
    \begin{array}{lll}
      \mx{ss1}{1} &:=& \elab{(\locflip \fplus \mesg{s2} \fplus \mesg{s3})}{1} \\ 
      \mx{ss2}{2} &:=& \elab{(\mesg{s1} \fplus \locflip \fplus \mesg{s3})}{2} \\
      \mx{ss3}{3} &:=& \elab{(\mesg{s1} \fplus \mesg{s2} \fplus \locflip)}{3} 
    \end{array}
    $$
  \end{block}

  Additive homorphism-- each $\mx{ssi}{i}$ for $i \in \setit{1,2,3}$ is a share:
  \begin{itemize}
  \item \emph{Algebraically}, $\mx{ss1}{1} + \mx{ss1}{2} + \mx{ss1}{3} \eop \sx{1}{1} + \sx{2}{2} + \sx{3}{3}$ 
  \item \emph{Probabilistically}, protocol step is noninteractive and each $\mx{ssi}{i}$ remains in a uniform distribution.
  \end{itemize}

\end{frame}

\begin{comment}
\begin{frame}{Example contd.: Output}

  \begin{itemize}
    \item Any $\mx{ss1}{*},\mx{ss2}{*},\mx{ss3}{*}$ such that:
    $$
    \mx{ss1}{*} + \ \mx{ss2}{*} + \ \mx{ss3}{*} \eop \sx{1}{1} + \sx{2}{2}  + \sx{3}{3}
    $$
    are \cemph{equally likely}.
  \end{itemize}

  \begin{block}{Client Outputs}
    $$
    \begin{array}{lll}
      \out{1} &:=& \elab{(\mesg{ss1} \fplus \mesg{ss2} + \mesg{ss3})}{1}\\
      \out{2} &:=& \elab{(\mesg{ss1} \fplus \mesg{ss2} + \mesg{ss3})}{2}\\
      \out{3} &:=& \elab{(\mesg{ss1} \fplus \mesg{ss2} + \mesg{ss3})}{3}
    \end{array}
    $$
  \end{block}

  %\begin{exampleblock}{Example Bad Protocol}
  %  If 1 is honest and accidentally broadcasts its local share:
  %  $$
  %  \mx{leak}{*} := \elab{\locflip}{1}
  %  $$
  %  conditioning on adversarial views now yields probability 1 for
  %  either $\sx{1}{1} = 0 \wedge \sx{2}{2} = 1$ or
  %  $\sx{1}{1} = 1 \wedge \sx{2}{2} = 0$.
  %\end{exampleblock}

\end{frame}
\end{comment}

\section{Using SMT}

\begin{frame}{cvc5 Finite Fields Theory}

  The cvc5 Finite Fields Theory supports automatate satisfiability checking of equational
  constraint definitions $\eqs$ over prime fields $\mathcal{F}_{p}$.

  \medskip

  Overture protocols $\prog$ have a straightforward interpretation as SMFF constraints $\toeq{\prog}$ , e.g.,
  the first two lines of $\prog_{\mathit{sum3}}$ translate as:
  $$
  %\toeq{
  %  \begin{array}{c}
  %  \begin{array}{lll}
  %    \elab{\mesg{s1}}{2} &:=& \elab{(\secret{1} \fminus \locflip \fminus \flip{x})}{1} \\ 
  %    \elab{\mesg{s1}}{3} &:=& \elab{\flip{x}}{1} 
  %  \end{array}\\
  %  = \\
    \begin{array}{lll}
      \elab{\mesg{s1}}{2} &\eop& \elab{\secret{1}}{1} \fminus \elab{\locflip}{1} \fminus \elab{\flip{x}}{1}\quad \wedge \\ 
      \elab{\mesg{s1}}{3} &\eop& \elab{\flip{x}}{1} 
    \end{array}
    %\end{array}
    %}
 $$

 Verifiable \emph{entailment} $\eqs_1 \models \eqs_2$ means that any model of $\eqs_1$ is a model
 of $\eqs_2$. Models of $\prog$ correspond exactly to valid runs of $\prog$...

\end{frame}

\section{SMT Boosting: Correctness}

\begin{frame}{Verifying Protocol Correctness}

  Entailment is a mechanism for verifying protocol properties. If an interpretation
  of a protocol entails a property, it holds in all protocol runs. 
$$
\begin{array}{c}
  \toeq{\sumthree} \wedge \ttt{out} \eop \mx{ss1}{1} + \mx{ss1}{2} + \mx{ss1}{3}\\ \models\\
  \ttt{out} \eop \sx{1}{1} \fplus \sx{2}{2} \fplus \sx{3}{3}
\end{array}
$$

Correctness of subroutines (e.g., addition and multiplication gates in encrypted circuits)
or the full protocol wrt $\idealf$ can be checked automatically.
\end{frame}

\section{SMT Boosting: Confidentiality}

\begin{frame}{Verifying Confidentiality}

$$    
  \inferrule[Encode]
  {\eqs \models \phi \eop \phi' \oplus \rx{w}{\cid} \\
   \oplus \in \{ \fplus,\fminus \}\\
   \eqj{R}{\eqs}{\phi'}{\ty}}
  {\eqj{R;\{ \rx{w}{\cid} \}}{\eqs}{\phi}{\setit{\cty{\rx{w}{\cid}}{\ty}}}}
$$

\end{frame}

\begin{frame}[fragile]{Modular Typing in Prelude}
  \small{
  \begin{verbatimtab}
SHARED(x) { (m[x]@1 + m[x]@2) }

andtablegmw(x, y, z) {
  let r11 = (r[z] + (m[x] + 1) * (m[y] + 1)) in
  ... // elided
  { r1 = r11; r2 = r10; r3 = r01; r4 = r00 }
}

andgmw(z:string, x:string, y:string) {
  let tab = andtablegmw(x,y,z) in
  m[z]@2 := OT4((m[x], m[y])@2, tab.r4, tab.r3, tab.r2, tab.r1)@1;
  m[z]@1 := r[z]@1
}
postcondition: ( m[z]@1 == r[z]@1 AND
                 m[z]@2 == (SHARED(x) * SHARED(y)) + r[z]@1) ) 
  \end{verbatimtab}
  }
\end{frame}

\section{Conclusion and Future Work}

\logo{
\includegraphics[width=3cm]{UVM_Logo_Primary_Horiz_G.png}
}

\begin{frame}{Future Work}

  \begin{itemize}
  \item Generalize verification to arbitrary prime fields.
  \item Automated type system to eliminate manual proof elements.
  \item Support for concurrency.
  \end{itemize}

\end{frame}

\end{document}