### 2025-05-10 esop25/new-cpj.tex
checkpoint

\renewcommand{\unity}[2]{U(#1, #2)}
\renewcommand{\ty}{T}
\newcommand{\eset}{D}
\newcommand{\dty}[2]{\langle #1, #2 \rangle}
\newcommand{\sharet}[1]{\mathrm{shares}(#1)}
\newcommand{\cvec}[1]{[#1]}
\newcommand{\secopent}{\mathrm{secopen}}
\newcommand{\declasst}[1]{\mathrm{declass}(#1)}
\renewcommand{\view}{\mathit{view}}
\newcommand{\revealed}{\mathit{revealed}}
\newcommand{\sty}{\varsigma}
\newcommand{\noise}{\mathrm{noise}}
\newcommand{\owners}{\mathrm{owners}}
\newcommand{\federation}{I_{\mathit{fed}}}
\newcommand{\tjoin}{\bowtie}

\section{Declassification Typing}

$$
\begin{array}{rcl@{\hspace{4mm}}r}
\eset &\in& 2^{\phi} & \gdesc{Dependency Approximations}\\ 
\ty &::=& \dty{R}{\eset} & \gdesc{Probabilistic Types} \\
\Gamma &::=& \varnothing \mid \Gamma; x : \ty & \gdesc{Type Environments}
\end{array}
$$

The types of expressions $\elab{\be}{\cid}$ reflect their dependencies which
may include variables from clients beside $\cid$, so typings are in terms
of interpretations $\toeq{\elab{\be}{\cid}}$. Types are of the form $\dty{R}{\eset}$,
where $\eset$ is a set of expressions $\phi$, and intuitively the typing
$\toeq{\elab{\be}{\cid}} : \dty{R}{\eset}$ means that the marginal distribution
of its variable dependencies is equivalent to the (uniform) marginal distribution of
variables in $R$. If $R$ is empty this means that the distribution is dependent
on the marginal distribution of variables in $\eset$. So for example:
\begin{mathpar}
  \sx{w}{\cid} : \dty{\varnothing}{\setit{\sx{w}{\cid}}}

  \rx{w}{\cid} : \dty{\setit{\rx{w}{\cid}}}{\varnothing}

  \sx{w}{\cid} - \rx{w}{\cid} : \dty{\setit{\rx{w}{\cid}}}{\setit{\sx{w}{\cid}}}
\end{mathpar}
The typing rules are an approximation of dependencies, so wherever dependencies
may be revealed, the system (conservatively) reflects this. So for example:
\begin{mathpar}
  \sx{w}{\cid} \ftimes \rx{w}{\cid} : \dty{\varnothing}{\setit{\rx{w}{\cid}, \sx{w}{\cid}}}
\end{mathpar}
\subsection{Primitive Typing Rules}

The primitive typing rules are precise and expressive and reflect no particular
implementation strategies, they just track dependencies as reflected in types.
The most important rules are for binary operators, where dependencies merge
in interesting ways especially with summation. In particular, if two
randomized values are summed, they remain randomized modulo disjointness
of randomness. Otherwise, overlapping randomness is merged into the dependencies.
For example:
\begin{eqnarray*}
& (\sx{w}{\cid_1} - \rx{w}{\cid_1}) + (\sx{w}{\cid_2} - \rx{w}{\cid_2}) : \\
& \dty{\setit{\rx{w}{\cid_1},\rx{w}{\cid_2}}}{\setit{\sx{w}{\cid_2},\sx{w}{\cid_1}}}
\end{eqnarray*}
but:
$$
(\sx{w}{\cid_1} - \rx{w}{\cid_1}) + \rx{w}{\cid_1} :
\dty{\varnothing}{\setit{\rx{w}{\cid_1},\sx{w}{\cid_1}}}
$$

\begin{definition}[Type Join]
  Given $\dty{R_1}{\eset_1}$ and $\dty{R_2}{\eset_2}$, define
  $\dty{R_1}{\eset_1} \tjoin \dty{R_2}{\eset_2} \defeq \dty{R}{\eset}$ where
  $R = (R_1 - (R_2 \cup \eset_2) \cup (R_2 - (R_1 \cup \eset_1))$ 
  and $\eset = (R_1 \cup R_2 \cup \eset_1 \cup \eset_2) - R $.
  %\begin{eqnarray*}
  %  R &\defeq& (R_1 - (R_2 \cup \eset_2) \cup (R_2 - (R_1 \cup \eset_1))\\
  %  \eset &\defeq& (R_1 \cup R_2 \cup \eset_1 \cup \eset_2) - R 
  %\end{eqnarray*}
\end{definition}

\begin{mathpar}
  \inferrule[Mesg]
            {}
            {\Gamma, \eqs \vdash \mx{w}{\cid} : \Gamma(\mx{w}{\cid})}

  \inferrule[Rand]
            {}
            {\Gamma, \eqs \vdash \rx{w}{\cid} : \dty{\setit{\rx{w}{\cid}}}{\varnothing}}

  \inferrule[Ideal]
            {}
            {\Gamma, \eqs \vdash f(\sx{w_1}{\cid_1},\ldots,\sx{w_n}{\cid_n}) :
              \dty{\varnothing}{\setit{f(\sx{w_1}{\cid_1},\ldots,\sx{w_n}{\cid_n})}}}

  \inferrule[Invert]
            {\Gamma, \eqs \vdash \phi : \ty}
            {\Gamma, \eqs \vdash -\phi : \ty}

  \inferrule[Join]
      {\Gamma, \eqs \vdash \phi_1 : \dty{R_1}{\eset_1}\\
       \Gamma, \eqs \vdash \phi_2 : \dty{R_2}{\eset_2}}%\\
        %R \defeq (R_1 - (R_2 \cup \eset_2) \cup (R_2 - (R_1 \cup \eset_1)) \\ 
        %\eset \defeq (R_1 \cup R_2 \cup \eset_1 \cup \eset_2) - R }
      {\Gamma, \eqs \vdash \phi_1 \fplus \phi_2 : \dty{R_1}{\eset_1} \tjoin \dty{R_2}{\eset_2}}

  \inferrule[Smear]
      {\Gamma, \eqs \vdash \phi_1 : \dty{R_1}{\eset_1}\\
       \Gamma, \eqs \vdash \phi_2 : \dty{R_2}{\eset_2}}
      {\Gamma, \eqs \vdash \phi_1 \bop \phi_2 : \dty{\varnothing}{R_1 \cup R_2 \cup \eset_1 \cup \eset_2}}

  \inferrule[Eq]
            {\eqs \models \phi \eop \phi' \\ \Gamma, \eqs \vdash \phi' : \ty}
            {\Gamma, \eqs \vdash \phi : \ty}          
\end{mathpar}

\begin{mathpar}
  \inferrule[Mesg]
      {\Gamma, \eqs \vdash \toeq{\elab{e}{\cid}} : \ty}
      {\Gamma, \eqs \vdash \xassign{x}{e}{\cid} : (\Gamma;x:\ty)}

  \inferrule[Seq]
      {\Gamma_1,\eqs \vdash \cmd_1 : \Gamma_2 \\
        \Gamma_2,\eqs \vdash \cmd_2 : \Gamma_3}
      {\Gamma_1, \eqs \vdash \cmd_1;\cmd_2 : \Gamma_3}
\end{mathpar}

\begin{definition}[Views]
  Given $\prog, \eqs, \Gamma$ with $\toeq{\prog} \models \eqs$ and $\varnothing, \eqs \vdash \prog : \Gamma$.
  Define:
  $$
  \view(\Gamma,\eqs,X) \defeq \bigcup_{X' \in (\pow(X) - \varnothing)} \ty
     \qquad \text{where}\ \Gamma,\eqs \vdash \Sigma_{x \in X'} x : \ty
  $$
\end{definition}

\begin{definition}[Revelation]
  $$
  \begin{array}{ll}
    \revealed(\ty,H) \defeq & \setit{f(\sx{w_1}{\cid_1},\ldots,\sx{w_n}{\cid_n}) \mid  \\
      & \qquad \dty{\varnothing}{\eset} \in \ty\ \wedge\\
      & \qquad f(\sx{w_1}{\cid_1},\ldots,\sx{w_n}{\cid_n}) \in \eset\ \wedge \\
      & \qquad \exists 1 \le i \le n . \cid_i \in H} 
  \end{array}
  $$
\end{definition}

\begin{theorem}[Security Typing]
  Given $\prog, \eqs, \Gamma$ with $\toeq{\prog} \models \eqs$ and $\varnothing, \eqs \vdash \prog : \Gamma$.
  Then $\prog$ satisfies {NIMO} for ideal functionality $\idealf$ if for all $H,C$, we have:
  $$
  \revealed(\view(\Gamma,\eqs,\vars(\prog)_C),H) = \setit{\idealf}
  $$
\end{theorem}

\subsection{Strategy Typings}

$$
\inferrule
    {\Gamma, \eqs \vdash \phi : \dty{R}{\eset} \\ R - R_I \neq \varnothing}
    {\Gamma, \eqs \vdash \phi : \noise(R,I)}
$$

\begin{mathpar}
  \inferrule[Sharing]
      {
        \eqs \models \mx{w_1}{\cid_1} \fplus \cdots \fplus \mx{w_n}{\cid_n} \eop \phi \\
        \\
        \exists R_2\ .\ (\forall M \in \pow(\setit{\mx{w_1}{\cid_1},\ldots,\mx{w_n}{\cid_n}})\ .\ 0 < |M| < n \implies \\
          \qquad \Gamma, \eqs \vdash \Sigma_{x \in M} x : \noise(R,\owners(M)) \ \wedge \ R \subseteq R_2)
      }   
      {\Gamma,\eqs \vdash R_1,\Delta \leadsto R_1;R_2, \Delta; \cvec{\mx{w_1}{\cid_1},\ldots,\mx{w_n}{\cid_n}} : \sharet{\phi}}

  \inferrule[Homomorphic Addition]
      {
       \Delta(\cvec{\mx{w^1_1}{\cid_1},\ldots,\mx{w^1_n}{\cid_n}}) = \sharet{\phi_1} \\
       \Delta(\cvec{\mx{w^2_1}{\cid_1},\ldots,\mx{w^2_n}{\cid_n}}) = \sharet{\phi_2} \\
       \eqs \models \mx{w_1}{\cid_1} \eop \mx{w^1_1}{\cid_1} \fplus \mx{w^2_1}{\cid_1} 
       \ \ldots\ 
       \eqs \models \mx{w_n}{\cid_n} \eop \mx{w^1_n}{\cid_n} \fplus \mx{w^2_n}{\cid_n} 
      }
      {
        \Gamma, \eqs \vdash R,\Delta \leadsto R,\Delta;\cvec{\mx{w_1}{\cid_1},\ldots,\mx{w_n}{\cid_n}} : \sharet{\phi_1 \fplus \phi_2}
      }

 \inferrule[Secure Opening]
      {
        \Delta(\cvec{\mx{w_1}{\cid_1},\ldots,\mx{w_n}{\cid_n}}) = \sharet{\phi} \\
        \eqs \models \px{w_1} \eop \mx{w_1}{\cid_1}
        \ \ldots\
        \eqs \models \px{w_n} \eop \mx{w_n}{\cid_n}\\
        \Gamma,\eqs \vdash \phi : \noise(R_2, \federation)
      }
      {
        \Gamma, \eqs \vdash R_1;\Delta \leadsto R_1;R_2, \Delta; \cvec{\px{w_1},\ldots,\px{w_n}} : \secopent
      }

 \inferrule[Declassification]
      {
        \Delta(\cvec{\mx{w_1}{\cid_1},\ldots,\mx{w_n}{\cid_n}}) =  \sharet{\phi} \\
        \eqs \models \px{w_1} \eop \mx{w_1}{\cid_1}
        \ \ldots\
        \eqs \models \px{w_n} \eop \mx{w_n}{\cid_n}
      }
      {
        \Gamma, \eqs \vdash R, \Delta \leadsto R, \Delta;\cvec{\px{w_1},\ldots,\px{w_n}} : \declasst{\phi}
      }
\end{mathpar}

\begin{theorem}
  Given $\prog, \eqs, \Gamma$ with $\toeq{\prog} \models \eqs$ and $\varnothing, \eqs \vdash \prog : \Gamma$.
  Suppose also $\Gamma,\eqs \vdash \varnothing,\varnothing \Rightarrow \Delta$, and define:
  $$\eset \defeq \{ \phi \mid \exists \nu \in \dom(\Delta) . \Delta(\nu) \mapsto \declasst{\phi} \}$$
  Let $\iov(\prog) = S \cup M \cup P$. 
  Then for any $H,C$, if for all $x \in (M \cup P)_C$ there exists
  $\nu \in \dom(\Delta)$ with $x \in \nu$, we have:
  $$
  D = \revealed(\view(\Gamma,\eqs,\vars(\prog)_C),H)
  $$
\end{theorem}

\begin{lemma}
  Given $\prog, \eqs, \Gamma$ with $\toeq{\prog} \models \eqs$ and $\varnothing, \eqs \vdash \prog : \Gamma$.
  Suppose also $\Gamma,\eqs \vdash \varnothing,\varnothing \Rightarrow \Delta$, and define:
  $$\eset = \{ \phi \mid \exists \nu \in \dom(\Delta) . \Delta(\nu) \mapsto \declasst{\phi} \}$$
  Then for any $H,C$, with:
  $$
  X \defeq \{ x \mid x \in \vars(\prog)_C \wedge \neg \exists \nu \in \dom(\Delta) . x \in \nu \}
  $$
  we have:
  $$
  \eset \cup  \revealed(\view(\Gamma,\eqs,X),H) = \revealed(\view(\Gamma,\eqs,\vars(\prog)_C),H)
  $$
\end{lemma}