### 2025-06-05 esop25/sharety.tex
checkpoint

\section{Share Types}

\newcommand{\sharety}{\mathrm{shares}}
\newcommand{\pubty}{\mathrm{public}}
\newcommand{\genshares}{\mathrm{genshares}}
\newcommand{\recon}{\mathrm{recon}}
\newcommand{\xvec}{\nu}

\begin{eqnarray*}
  \genshares(\phi,y,\cid) &\defeq& \setit{\rx{y}{\cid}}, [\phi - \rx{y}{\cid}, \rx{y}{\cid}]\\
  \recon(\phi_1,\phi_2) &\defeq& \phi_1 \fplus \phi_2
\end{eqnarray*}

\begin{mathpar}
  \inferrule[Share]
      {\eqs \models \genshares(\phi_1,y,\cid) \eop R_2,[\mx{z}{\cid},\phi_2] }
      {\Gamma,R_1,\eqs \vdash [\mx{z}{\cid}] : R_1;R_2, \Gamma; [\mx{z}{\cid},\phi_2] : \sharety}

  \inferrule[Open]
      {\Gamma,R,\eqs \vdash [\px{y},\px{z}] : \sharety}
      {\Gamma,R,\eqs \vdash [\px{y},\px{z}] : R, \Gamma; [\recon(\px{y},\px{z})] : \pubty}

  \inferrule[Seq]
      {\Gamma_1,\eqs,R_1 \vdash \xvec_1 : R_2,\Gamma_2 \\
        \Gamma_2,\eqs,R_2 \vdash \xvec_2 : R_3,\Gamma_3}
      {\Gamma_1, \eqs,R_1 \vdash \xvec_1\xvec_2 : R_3,\Gamma_3}
\end{mathpar}

\begin{mathpar}
  \inferrule[Hyp]
      {}
      {\Gamma,\eqs \vdash \xvec : \Gamma(\xvec)}

  \inferrule[HEAdd]
      {\Gamma,\eqs \vdash [\phi^1_1,\phi^1_2] : \sharety\\
        \Gamma,\eqs \vdash [\phi^2_1,\phi^2_2] : \sharety}
      {\Gamma,\eqs \vdash [\phi^1_1\fplus\phi^2_1,\phi^1_2\fplus\phi^2_2] : \sharety}

  \inferrule[HEAddPub]
      {\Gamma,\eqs \vdash [\phi] : \pubty\\
        \Gamma,\eqs \vdash [\phi_1,\phi_2] : \sharety}
      {\Gamma,\eqs \vdash [\phi_1\fplus\phi,\phi_2] : \sharety}

  \inferrule[HEMultPub]
      {\Gamma,\eqs \vdash [\phi] : \pubty\\
        \Gamma,\eqs \vdash [\phi_1,\phi_2] : \sharety}
      {\Gamma,\eqs \vdash [\phi_1\ftimes\phi,\phi_2\ftimes\phi] : \sharety}

  \inferrule[PubOp]
      {\Gamma,\eqs \vdash [\phi_1] : \pubty\\
        \Gamma,\eqs \vdash [\phi_2] : \pubty}
      {\Gamma,\eqs \vdash [\phi_1\bop\phi_2] : \pubty}
\end{mathpar}