### 2025-11-14 toplas26/frontmatter.tex
checkpoint

\title{SMT-Boosted Security Types for Low-Level MPC}

\author{Christian Skalka}
\affiliation{%
  \institution{University of Vermont}
  \city{Burlington, Vermont}
  \country{USA}}
\email{ceskalka@uvm.edu}

\author{Joseph P.~Near}
\affiliation{%
  \institution{University of Vermont}
  \city{Burlington, Vermont}
  \country{USA}}
\email{jnear@uvm.edu}

\author{Steven Baldasty}
\affiliation{%
  \institution{University of Vermont}
  \city{Burlington, Vermont}
  \country{USA}}
\email{sbaldast@uvm.edu}

\author{Yeonhee Yeh}
\affiliation{%
  \institution{University of Vermont}
  \city{Burlington, Vermont}
  \country{USA}}
\email{yyeh@uvm.edu}

\begin{abstract}
  Secure Multi-Party Computation (MPC) is an important enabling
  technology for data privacy in modern distributed applications. We
  develop a new type theory to automatically enforce correctness,
  confidentiality, and integrity properties of protocols written in
  the \emph{Prelude/Overture} language framework. Judgements in the
  type theory are predicated on SMT verifications in a theory of
  finite fields, which supports precise and efficient analysis. Our
  approach is automated, compositional, scalable, and generalizes to
  arbitrary prime fields.
\end{abstract}