### 2025-11-24 toplas26/frontmatter.tex
checkpoint

\title{SMT-Boosted Security Types and Correctness for Low-Level MPC}

\author{Christian Skalka}
\affiliation{%
  \institution{University of Vermont}
  \city{Burlington, Vermont}
  \country{USA}}
\email{ceskalka@uvm.edu}

\author{Steven Baldasty}
\affiliation{%
  \institution{University of Vermont}
  \city{Burlington, Vermont}
  \country{USA}}
\email{sbaldast@uvm.edu}

\author{Yeonhee Yeh}
\affiliation{%
  \institution{University of Vermont}
  \city{Burlington, Vermont}
  \country{USA}}
\email{yyeh@uvm.edu}

\author{Joseph P.~Near}
\affiliation{%
  \institution{University of Vermont}
  \city{Burlington, Vermont}
  \country{USA}}
\email{jnear@uvm.edu}

\begin{abstract}
  Secure Multi-Party Computation (MPC) is an important enabling
  technology for data privacy in modern distributed applications. We
  develop a new Hoare logic and type theory to automatically enforce
  correctness, confidentiality, and integrity properties of protocols
  written in the \emph{Prelude/Overture} language framework. Both
  definition and automation of our analyses leverages SMT verification
  in a theory of finite fields, supporting precision and
  efficiency. Our approach is compositional, scalable, and generalizes
  to arbitrary prime fields and secret sharing schemes.
\end{abstract}