### 2025-11-24 toplas26/conclusion.tex
checkpoint

\section{Conclusion and Future Work}

In this paper we have developed the $\minicat$ language for defining
low-level MPC protocols, and the $\metaprot$ metaprogramming language
for specifying $\minicat$ protocols compositionally and at a higher
level of abstraction. We have developed a Hoare logic based on
SMT-style verification for $\minicat$, as well as confidentiality and
integrity type systems that enforce passive and malicious security in
the real/ideal model. These establish a foundation for an automated
Hoare logic and type system in $\metaprot$ that is compositional,
scalable, and generalizes to arbitrary prime fields and secret sharing
schemes.

Our automated $\metaprot$ Hoare logic leverages Satisfiability
Modulo Finite Fields for efficiency, scalability, and generalizability
to arbitrary prime fields. We developed examples of real binary and
arithmetic circuit protocols, illustrated the applicability of
our analyses to enforce confidentiality and integrity properties,
and evaluated performance of our implementation on these examples.